Cominuscule subvarieties of flag varieties

### 2025-01-04 ms.tex
Add files via upload

\usepackage{lmodern}
\usepackage{cfr-lm}
\usepackage[T2A,T1]{fontenc}
\usepackage[utf8]{inputenx}
\usepackage{siunitx}
\usepackage{xparse}
\usepackage{standalone}
\usepackage{mathtext}
\usepackage[english]{babel}
\ifdefined\hypersetup%
\else%
\usepackage[pdftex]{hyperref}%
\fi%
\hypersetup{
  colorlinks   = true,  %Colours links instead of ugly boxes
  urlcolor     = black, %Colour for external hyperlinks
  linkcolor    = black, %Colour of internal links
  citecolor    = black  %Colour of citations
}
\usepackage[kerning=true,tracking=true]{microtype}
\usepackage{xspace}
\usepackage{amsfonts}
\usepackage{verbatim}
\usepackage{amssymb}
\usepackage{mathtools}
\usepackage{mathabx}
\usepackage{braket}
\usepackage{cool}
\usepackage{varioref}
\usepackage{xstring}
\usepackage{array}
\usepackage{ragged2e}
\usepackage{longtable}
\usepackage[longtable]{multirow}
\usepackage{booktabs}
\usepackage{enumitem}
\usepackage{varwidth}
\usepackage{tensor}
\usepackage{scalerel}
\usepackage{lie-hasse}
\tikzset{
	background rectangle/.style={
		shade,
		top color=olive!20,
		bottom color=white,
		draw=olive!15,
		very thick,
		rounded corners}}
\usepackage{rank-2-roots}
\usepackage{tikz-cd}
\usepackage{colortbl}
\newtheorem{theorem}{Theorem}
\newtheorem{corollary}{Corollary}
\newtheorem{lemma}{Lemma}
\newtheorem{proposition}{Proposition}
\theoremstyle{remark}
\newtheorem{conjecture}{Conjecture}
\newtheorem{definition}{Definition}
\newtheorem{example}{Example}
\newtheorem{remark}{Remark}
\newtheorem{Comment}{Comment}
\newcounter{remarkCounter}
\setcounter{remarkCounter}{1}
\NewDocumentCommand\todo{m}
{%
    \par\noindent{\textcolor{blue}{\textbf{Comment \#\arabic{remarkCounter}. } #1}}
    \par
    \stepcounter{remarkCounter}
}%
\NewDocumentCommand\finish{}{\todo{Finish this.}}

@@

\NewDocumentCommand\drawroots{mm}%
{%
\begin{tikzpicture}[baseline=-.5]
\begin{rootSystem}{#1}
\roots
\parabolic{#2}
\parabolicgrading
\simpleroots
\end{rootSystem}
\end{tikzpicture}
}%
\NewDocumentCommand\mx{m}{{#1}^+}
\NewDocumentCommand\mn{m}{{#1}^-}
\NewDocumentCommand\cp{m}{{#1}^{\degree}}
\makeatletter
\newcommand*{\KillingForm}[2]%
{%%
\ensuremath{%%%
%#1\ifnum\pdf@strcmp{#1}{#2}=\z@ ^2\else\cdot#2\fi
#1\cdot#2
}%%%
}%%
\makeatother
\NewDocumentCommand\KillingSquare{m}%
{%
\ensuremath{#1^2}%
}%
\NewDocumentCommand\om{m}{\omega^{#1}}

@@

@@

@@

@@

Flag varieties have few regular maps between them \cite{bakshi2023morphisms,kumar2023nonexistence,naldi2022morphisms,occhetta2023morphisms,Sierra:2021,Tango:1974,Tango:1976}, hence few flag subvarieties, so these subvarieties are surprising, and we expect them to be important.

@@

As in the image of Aten's rays, pick a point \(p_0\) and a line \(\ell_0\) in projective space \(\Proj{2}\), with \(p_0\) not lying on \(\ell_0\).

@@

@@

V_2\subseteq W_2, V_2\oplus V_3\subseteq W_3, \dots, V_2\oplus\dots V_{k-1}\subseteq W_{k-1},

@@

\dim ((V_2\oplus\dots V_{k-1})\cap W_{k-1})\ge n_1+\dots+n_{k-1},

@@

A \emph{flag variety} \((X,G)\), also called a \emph{generalized flag variety} or a \emph{rational homogeneous variety}, is a complex projective variety \(X\) acted on transitively and holomorphically by a connected complex semisimple Lie group \(G\) \cite{Knapp:2002} p. 325.

@@

(A subgroup of \(G\) is \emph{parabolic} if it is the stabilizer of a point of a flag variety \((X,G)\) of \(G\), hence the use of the letter \(P\).)

Denote the Lie algebras of \(P\subseteq G\) by \(\LieP\subset\LieG\). 
One can select a Cartan subgroup of \(G\) lying inside \(P\), whose positive root spaces all lie in \(\LieP\).
A simple root \(\aa\) is \(P\)-\emph{compact} (\emph{compact} if \(P\) is understood) if the root space of \(-\aa\) belongs to the Lie algebra of \(P\).
Each flag variety is determined uniquely, up to finite central extension of \(G\) and up to isomorphism, by the Dynkin diagram of \(G\) decorated with \(\dynkin{A}{*}\) on each compact simple root and \(\dynkin{A}{x}\) on each noncompact root \cite{Borel:1991} p. 197 Proposition 14.18.
\subsection{Reducible flag varieties}\label{subsec:reducible}
The center of any complex semisimple Lie group \(G\) lies in every maximal torus, so in every Cartan subgroup \cite{Borel:1991} p. 220, so in every parabolic subgroup, so acts trivially on every flag variety.
An \emph{irreducible} flag variety is a flag variety \((X,G)\) with \(G\) a simple Lie group, and with only the center of \(G\) acting trivially.
Every flag variety \((X,G)\), up to finite central extension of \(G\), admits a factorization 
\begin{align*}
X&=X_0 \times X_1 \times X_2 \times \dots \times X_s, \\
G&=G_0 \times G_1 \times G_2 \times \dots \times G_s, 
\end{align*}
into irreducible flag varieties \(\pr{X_i, G_i}\), \(i>0\), and a point \(X_0=\set{x_0}\), unique up to permutation of the \((X_i,G_i)\) for \(i>0\) and isomorphism.
The flag variety \((X,G)\) is effective if and only if all \((X_i,G_i)\) are effective, i.e. if and only if \(G_0=\set{1}\) is trivial and \(G_1,\dots,G_s\) are in adjoint form.
\subsection{Cominuscule varieties}
A flag variety is \emph{cominuscule} if \(\LieG/\LieP=T_{x_0} X\) is a sum of irreducible complex algebraic \(P\)-modules.
This occurs just when there is a compact subgroup \(K\subseteq G\) so that \(\pr{X,K}\) is a compact Hermitian symmetric space \cite{Kostant:1961} p. 379 Proposition 8.2, \cite{Baston/Eastwood:1989} p. 26.
Some authors prefer the term \emph{compact Hermitian symmetric space}, \emph{cominuscule Grassmannian}, or \emph{generalized Grassmannian} to \emph{cominuscule variety}.
Every effective cominuscule variety is a product of the following irreducible effective cominuscule varieties \cite{Kobayashi/Nagano:1964} theorem 1 p. 401:\par\noindent{}%
%%\input{hss}
\begingroup
\small
\begin{longtabl}{@{}%
>{\columncolor[gray]{.9}$}r<{$}%
>{\columncolor[gray]{.93}}p{2.6cm}%
>{\columncolor[gray]{.9}$}r<{$}%
>{\columncolor[gray]{.93}}l@{}}
\toprule
G&\(G/P\)&\operatorname{dim}&\text{description}\\
\midrule
\endfirsthead
\toprule
G&\(G/P\)&\operatorname{dim}&\text{description}\\
\midrule
\endhead
\bottomrule
\endfoot
\bottomrule
\endlastfoot
A_r&\dynkin[labels={,,,k,,,}]{A}{**.*x*.**}&k(r+1-k)&Grassmannian of $k$-planes in $\C[r+1]$
\\
B_r&\dynkin[parabolic=1]{B}{}&2r-1&quadric hypersurface in $\Proj{2r}$\\
C_r&\dynkin[parabolic=16]{C}{}&\frac{r(r+1)}{2}&space of Lagrangian $r$-planes in $\C[2r]$\\
D_r&\dynkin[parabolic=1]{D}{}&2r-2&quadric hypersurface in $\Proj{2r-1}$\\
D_r&\dynkin[parabolic=32]{D}{}&\frac{r(r-1)}{2}& %one component of the variety of 
space of null $r$-planes in $\C[2r]$ \\
E_6&\dynkin[parabolic=1]{E}{6}&16&complexified octave projective plane\\
E_7 &\dynkin[parabolic=64]{E}{7}&27&space of null octave \(3\)-planes in octave \(6\)-space
\end{longtabl}
\endgroup
%% end input
\subsection{Structure of linear algebraic groups}
A complex linear map is \emph{unipotent} if its only eigenvalue is \(1\).
A subgroup of a linear algebraic group is \emph{unipotent} if it consists of unipotent linear maps.
Every complex linear algebraic group \(G\) has a \emph{unipotent radical}, the unique maximal unipotent normal subgroup, which is a closed complex linear algebraic subgroup \cite{Borel:1991} p. 85 Theorem 4.5, p. 86 Theorem 4.7, p. 157 11.21.

A complex linear algebraic group is \emph{reductive} if it contains a Zariski dense compact subgroup \cite{Knapp:2002} p. 382 Proposition 7.15.
Every complex linear algebraic group \(G\) has a \emph{reductive Levi factor}, i.e. a maximally reductive complex linear algebraic subgroup,  unique up to conjugacy, so that \(G\) is a semidirect product of its reductive Levi factor and its unipotent radical \cite{Borel:1991} p. 158 11.22.
Any Cartan subgroup is therefore a subgroup of the Levi factor, after perhaps a conjugacy.
Every compact subgroup lies in a maximal compact subgroup, which is unique up to conjugacy, so lies in the Levi factor up to conjugacy \cite{Hilgert.Neeb:2012} p. 531 Theorem 14.1.3.
After perhaps extending by some finite group of order a power of \(2\), the Weyl group embeds in \(G\) as a finite subgroup \cite{Tits:1966}, hence compact, so this group lies in the Levi factor up to conjugacy.

@@

A Zariski closed subgroup \(P\subseteq G\) of a connected linear algebraic group \(G\) is \emph{parabolic} if \(X:=G/P\) is a projective variety, and this occurs just when \((X,G)\) is a flag variety, and just when \(P\) contains a maximal connected solvable subgroup \cite{Borel:1991} p. 148.

@@

\subsection{Opposite parabolic subgroups}

@@

(We can define such an automorphism explicitly as \(e_{\alpha}\mapsto -e_{-\alpha}\) on root vectors in a Chevalley basis; see section~\vref{subsubsection:ChevalleyBases}.)
Our automorphism sends \(P\) to its \emph{opposite} \(aP\) parabolic subgroup, so gives \(P\cap aP=\GZ\) and \(G_+\cap aG_+=\set{1}\) and an open subset of \(G\) consists of elements uniquely expressed as a product \(p,q\in P\times aG_+\mapsto pq\in G\) \cite{Cap/Slovak:2009} p. 294, \cite{Borel:1991} p. 198 Proposition 14.21.

@@

Let \(\breve{G}\defeq\left<Z,aZ\right> \subseteq G\) be the subgroup generated by \(Z\cup aZ\), \(\breve{P} \defeq \breve{G} \cap P\), \(\breve{X} \defeq \breve{G}/\breve{P}\).

@@

@@

@@

\section{Statement of the theorem}

@@

With \(\dynkin{A}{o}\) denoting a node which could be either a \(\dynkin{A}{x}\) or \(\dynkin{A}{*}\), the associated cominuscule subvarieties are:

@@

@@

\dynkinBrace[q]{7}{8}

@@

@@

@@

@@

\dynkinBrace[\ell\ge 2]{1}{4}

@@

\LastFlag{4}{B_r}{%

@@

\dynkinBrace[\ell\ge 3]{1}{4}

@@

@@

\Flag{\dynkin{C}{xo.ooo}}{\dynkin{A}{x}}%
\LastFlag{3}{C_r}{%

@@

\Flag{\dynkin[parabolic=1]{D}{}}{\dynkin[parabolic=1]{D}{}}%
\Flag{\dynkin[parabolic=32]{D}{}}{\dynkin[parabolic=32]{D}{}}%
\Flag{\dynkin[parabolic=16]{D}{}}{\dynkin[parabolic=16]{D}{}}%
\LastFlag{4}{D_r}{%
\begin{dynkinDiagram}{D}{o*.*xo.ooo}
\dynkinBrace[\ell\ge 2]{1}{4}

@@

\dynkinBrace[\ell-1]{1}{4}

@@

@@

@@

\Flag{\dynkin[ordering=Carter]{E}{*****xoo}}{\dynkin{A}{x*****}}

@@

\LastFlag{8}{E_8}{\dynkin[ordering=Carter]{E}{*******x}}{\dynkin[parabolic=1]{D}{8}}

@@

@@

For a flag variety \(X=G/P\), the root system is also \emph{\(P\)-graded} by \(\sum n_i\), but summing only over the noncompact simple roots.
The unipotent radical \(G_+\subseteq P\) has Lie algebra the sum of the root vectors of the positively graded roots; clearly this sum is a unipotent Lie algebra, while the zero graded roots are invariant under reflection in one another, giving the semisimple factor of the Levi factor.
A root is \emph{\(P\)-maximal} if it has highest grade in its irreducible factor.
Its root vector then lies inside the center of the unipotent radical of \(P\).
The \emph{box} is the set of \(P\)-maximal roots, terminology which roughly follows \cite{Buch.Chaput.Mihalcea.Perrin:2018}, \cite{Lam.Williams:2008} p. 57, by analogy with Young tableaux.

@@

\subsection{\texorpdfstring{Associated cominuscules in rank \(2\)}{Associated cominuscules in rank 2}}

@@

We pick simple roots in our root systems as:

@@

Each parabolic subgroup has Lie algebra consisting of the sum of the Cartan and the root spaces of those roots which lie on or on one side of a hyperplane, so that the compact roots are those on the hyperplane.

@@

We can always pick a basis of simple roots so that every simple root lies on the hyperplane (hence a compact simple root) or lies on the chosen side of the hyperplane (a noncompact root):

@@

(We will instead always pick our hyperplane to allow for the chosen bases of simple roots shown above.)
Grade the roots by height. 
Our hyperplane is always chosen so that we can see the heights, i.e. roots of a given height lie on a parallel hyperplane:

@@

>>>>

\usepackage{lmodern}
\usepackage{cfr-lm}
\usepackage[T2A,T1]{fontenc}
\usepackage[utf8]{inputenx}
\usepackage{lineno}
\linenumbers
\usepackage{siunitx}
\usepackage{xparse}
\usepackage{standalone}
\usepackage{mathtext}
\usepackage[english]{babel}
\usepackage{hyperref}%
\hypersetup{
  colorlinks   = true,  %Colours links instead of ugly boxes
  urlcolor     = black, %Colour for external hyperlinks
  linkcolor    = black, %Colour of internal links
  citecolor    = black  %Colour of citations
}
\usepackage[kerning=true,tracking=true]{microtype}
\usepackage{xspace}
\usepackage{amsfonts}
\usepackage{verbatim}
\usepackage{amssymb}
\usepackage{mathtools}
\usepackage{mathabx}
\usepackage{braket}
\usepackage{cool}
\usepackage{varioref}
\usepackage{xstring}
\usepackage{array}
\usepackage{ragged2e}
\usepackage{longtable}
\usepackage[longtable]{multirow}
\usepackage{booktabs}
\usepackage{enumitem}
\usepackage{varwidth}
\usepackage{tensor}
\usepackage{scalerel}
\usepackage{lie-hasse}
\tikzset{
	background rectangle/.style={
		shade,
		top color=olive!20,
		bottom color=white,
		draw=olive!15,
		very thick,
		rounded corners}}
\usepackage{rank-2-roots}
\usepackage{tikz-cd}
\usepackage{colortbl}
\newtheorem{theorem}{Theorem}
\newtheorem{corollary}{Corollary}
\newtheorem{lemma}{Lemma}
\newtheorem{proposition}{Proposition}
\theoremstyle{remark}
\newtheorem{conjecture}{Conjecture}
\newtheorem{definition}{Definition}
\newtheorem{example}{Example}
%\newtheorem{remark}{Remark}
%\newtheorem{Comment}{Comment}
%\newcounter{remarkCounter}
%\setcounter{remarkCounter}{1}
%\NewDocumentCommand\todo{m}
%{%
%    \par\noindent{\textcolor{blue}{\textbf{Comment \#\arabic{remarkCounter}. } #1}}
%    \par
%    \stepcounter{remarkCounter}
%}%
%\NewDocumentCommand\finish{}{\todo{Finish this.}}

@@

\NewDocumentCommand\drawroots{mm}%
{%
\begin{tikzpicture}[baseline=-.5]
\begin{rootSystem}{#1}
\roots
\parabolic{#2}
\parabolicgrading
\simpleroots
\end{rootSystem}
\end{tikzpicture}
}%
\NewDocumentCommand\mx{m}{{#1}^+}
\NewDocumentCommand\mn{m}{{#1}^-}
\NewDocumentCommand\cp{m}{{#1}^{\degree}}
\makeatletter
\newcommand*{\KillingForm}[2]%
{%%
\ensuremath{%%%
%#1\ifnum\pdf@strcmp{#1}{#2}=\z@ ^2\else\cdot#2\fi
#1\cdot#2
}%%%
}%%
\makeatother
\NewDocumentCommand\KillingSquare{m}%
{%
\ensuremath{#1^2}%
}%
\NewDocumentCommand\om{m}{\omega^{#1}}
\NewDocumentCommand\fl{om}{#2^{\IfValueTF{#1}{#1}{\scriptscriptstyle{\bullet}}}}
\NewDocumentCommand\gr{om}{#2_{\IfValueTF{#1}{#1}{\scriptscriptstyle{\bullet}}}}
% filtered vector space, Lie algebra, module, but remembering only the underlying integer grading.
\NewDocumentCommand\uZ{m}{{}^{\scriptscriptstyle\#}{#1}}
\NewDocumentCommand\odeg{m}{{#1}^{\scriptscriptstyle\#}}
\NewDocumentCommand\agr{som}%
{%
\operatorname{gr}_{\IfValueTF{#2}{#2}{\scriptscriptstyle{\bullet}}}\IfBooleanTF{#1}{(#3)}{\!#3}%
}%

@@

% opposite parabolic subgroup, subalgebra
\NewDocumentCommand\op{m}{#1^{\text{op}}}

@@

We study the tangent bundles of flag varieties.

@@

The Hasse diagrams as drawn by Claus Ringel \cite{Ringel2013} are very clear, so we will follow his conventions.

@@

Flag varieties have few regular maps between them \cite{bakshi2023morphisms,kumar2023nonexistence,naldi2022morphisms,occhetta2023morphisms,Sierra:2021,Tango:1974,Tango:1976}, hence few flag subvarieties, so these subvarieties are surprising.
Since cominiscule varieties are simpler than other flag varieties in many ways, we hope that these cominiscule varieties will shine light on their ambient flag varieties.
We give some evidence for their significance by demonstrating that the associated cominiscule subvariety of any irreducible flag variety is the unique submanifold of maximal symmetry group among all submanifolds satisfying a certain open condition on derivatives.

@@

As in the image of Aten's rays, pick a point \(p_0\) and a line \(\ell_0\) in the projective plane \(\Proj{2}\), with \(p_0\) not lying on \(\ell_0\).

@@

\draw[white,thick] (1,.75) -- (\x,\y);
\fill (1,.75) circle (2pt);
%\draw[white,opacity=.5,thick] ({\x+\r*cos(180+5*\i)},{\y+\r*sin(180+5*\i)}) -- ++({\R*cos(180+5*\i)},{\R*sin(180+5*\i)});

@@

V_2\subseteq W_2, V_2\oplus V_3\subseteq W_3, \dots, V_2\oplus\dots\oplus V_{k-1}\subseteq W_{k-1},

@@

\dim ((V_2\oplus\dots\oplus V_{k-1})\cap W_{k-1})\ge n_1+\dots+n_{k-1},

@@

A \emph{flag variety} \((X,G)\), also called a \emph{generalized flag variety} or a \emph{rational homogeneous variety}, is a complex projective variety \(X\) acted on transitively and holomorphically by a connected complex semisimple Lie group \(G\) \cite{Humphreys:1975} pp. 134--135.

@@

A subgroup of \(G\) is \emph{parabolic} if it is the stabilizer of a point of a flag variety \((X,G)\) of \(G\), hence the use of the letter \(P\).

Denote the Lie algebras of \(P\subseteq G\) by \(\LieP\subseteq\LieG\). 
One can select a Cartan subgroup of \(G\) lying inside \(P\), whose positive root spaces all lie in \(\LieP\).
A simple root \(\aa\) is \(P\)-\emph{compact} (\emph{compact} if \(P\) is understood) if the root space of \(-\aa\) belongs to the Lie algebra of \(P\).
Each flag variety is determined uniquely, up to finite central extension of \(G\) and up to isomorphism, by the Dynkin diagram of \(G\) decorated with \(\dynkin{A}{*}\) on each compact simple root and \(\dynkin{A}{x}\) on each noncompact simple root \cite{Borel:1991} p. 197 Proposition 14.18, \cite{Humphreys:1975} p. 197 Theorem${}'$.
\subsection{Reducible flag varieties}\label{subsec:reducible}
The center of any complex semisimple Lie group \(G\) lies in every maximal torus, so in every Cartan subgroup \cite{Borel:1991} p. 220, so in every parabolic subgroup, so acts trivially on every flag variety.
An \emph{irreducible} flag variety is a flag variety \((X,G)\) with \(G\) a simple Lie group, and with only the center of \(G\) acting trivially.
Every flag variety \((X,G)\), up to finite central extension of \(G\), admits a factorization 
X&=X_0 \times X_1 \times X_2 \times \dots \times X_s, \\
G&=G_0 \times G_1 \times G_2 \times \dots \times G_s, 
into irreducible flag varieties \(\pr{X_i, G_i}\), \(i>0\), and a point \(X_0=\set{x_0}\), unique up to permutation of the \((X_i,G_i)\) for \(i>0\) and isomorphism.
The flag variety \((X,G)\) is effective if and only if all \((X_i,G_i)\) are effective, i.e. if and only if \(G_0=\set{1}\) is trivial and \(G_1,\dots,G_s\) are in adjoint form, and then \((X,G)\) is precisely (not just up to finite central extension) \cite{Akhiezer:1995} p.~74 the product 
X&=X_1 \times X_2 \times \dots \times X_s, \\
G&=G_1 \times G_2 \times \dots \times G_s.
\subsection{Cominuscule varieties}
A flag variety is \emph{cominuscule} if \(\LieG/\LieP=T_{x_0} X\) is a sum of irreducible complex algebraic \(P\)-modules.
This occurs just when there is a compact subgroup \(K\subseteq G\) so that \(\pr{X,K}\) is a compact Hermitian symmetric space \cite{Kostant:1961} p. 379 Proposition 8.2, \cite{Baston/Eastwood:1989} p. 26.
Some authors prefer the term \emph{compact Hermitian symmetric space}, \emph{cominuscule Grassmannian}, or \emph{generalized Grassmannian} to \emph{cominuscule variety}.
Every effective cominuscule variety is a product of the following irreducible effective cominuscule varieties \cite{Kobayashi/Nagano:1964} theorem 1 p. 401:\par\noindent{}%
%%\input{hss}
\small
\begin{longtabl}{@{}%
>{\columncolor[gray]{.9}$}r<{$}%
>{\columncolor[gray]{.93}}p{2.6cm}%
>{\columncolor[gray]{.9}$}r<{$}%
>{\columncolor[gray]{.93}}l@{}}
A_r&\dynkin[labels={,,,k,,,}]{A}{**.*x*.**}&k(r+1-k)&Grassmannian of $k$-planes in $\C[r+1]$
C_r&\dynkin[parabolic=16]{C}{}&\frac{r(r+1)}{2}&space of Lagrangian $r$-planes in $\C[2r]$\\
D_r&\dynkin[parabolic=16]{D}{}&\frac{r(r-1)}{2}& %one component of the variety of 
component of pace of null $r$-planes in $\C[2r]$ \\
D_r&\dynkin[parabolic=32]{D}{}&\frac{r(r-1)}{2}& %one component of the variety of 
component of space of null $r$-planes in $\C[2r]$ \\
E_7 &\dynkin[parabolic=64]{E}{7}&27&space of null octave \(3\)-planes in octave \(6\)-space
\subsection{Structure of linear algebraic groups}
A complex linear map is \emph{unipotent} if its only eigenvalue is \(1\).
A subgroup of a linear algebraic group is \emph{unipotent} if it consists of unipotent linear maps.
Every complex linear algebraic group \(G\) has a \emph{unipotent radical}, the unique maximal unipotent normal subgroup, which is a closed complex linear algebraic subgroup \cite{Borel:1991} p. 85 Theorem 4.5, p. 86 Theorem 4.7, p. 157 11.21.

A complex linear algebraic group is \emph{reductive} if it contains a Zariski dense compact subgroup \cite{Vinberg:1994} p. 142 Theorem 2.7, Theorem 2.8.
Every complex linear algebraic group \(G\) has a \emph{reductive Levi factor}, i.e. a maximally reductive complex linear algebraic subgroup,  unique up to conjugacy, so that \(G\) is a semidirect product of its reductive Levi factor and its unipotent radical \cite{Borel:1991} p. 158 11.22.
Any Cartan subgroup is therefore a subgroup of the Levi factor, after perhaps a conjugacy.
Every compact subgroup lies in a maximal compact subgroup, which is unique up to conjugacy, so lies in the Levi factor up to conjugacy \cite{Hilgert.Neeb:2012} p. 531 Theorem 14.1.3.
After perhaps extending by some finite group of order a power of \(2\), the Weyl group embeds in \(G\) as a finite subgroup \cite{Tits:1966}, hence compact, so this group lies in the Levi factor up to conjugacy.

@@

A Zariski closed subgroup \(P\subseteq G\) of a connected linear algebraic group \(G\) is \emph{parabolic} if \(X:=G/P\) is a projective variety, and this occurs just when \((X,L)\) is a flag variety for a semisimple Levi factor \(L\subseteq G\), and this occurs just when \(P\) contains a Borel subgroup (i.e. a maximal connected solvable subgroup) \cite{Borel:1991} p. 148.

@@

\subsection{Opposite parabolic subgroups}\label{subsec:opposite.parabolic}
Two parabolic subgroups \(P,\op{P}\subseteq G\) of a complex semisimple Lie group are \emph{opposite} if \(P\cap\op{P}\) is a maximal reductive Levi subgroup of both \(P\) and \(\op{P}\).

@@

(We can define such an automorphism explicitly as \(e_{\alpha}\mapsto -e_{-\alpha}\) on root vectors in a Chevalley basis; see \S\vref{subsubsection:ChevalleyBases}.)
Our automorphism sends \(P\) to an opposite parabolic subgroup \(\op{P}:=aP\) with \(P\cap \op{P}=\GZ\).
Letting \(G_-:=aG_+\), \(G_+\cap G_-=\set{1}\).
An open subset of \(G\) consists of elements uniquely expressed as a product \(p,q\in P\times G_-\mapsto pq\in G\) \cite{Cap/Slovak:2009} p. 294, \cite{Borel:1991} p. 198 Proposition 14.21.
Every root system also has an automorphism, traditionally called \(w_0\), which belongs to the Weyl group and which interchanges the positive and negative roots of a root system \cite{Cap/Slovak:2009} p. 323-324; it is the unique element of the Weyl group of minimum length.
Note that \(w_0\) might not reverse the signs of simple roots \cite{Cap/Slovak:2009} p. 324.
The Weyl group lifts to a group of automorphisms of the Lie group \(G\), after perhaps extension by some finite group of order a power of \(2\).
We can use such an extension of \(w_0\) in place of \(a\) throughout this paper, as we will only need that \(a\) is an automorphism of a given root system which extends to an automorphism of \(G\) taking a given parabolic subgroup to an opposite.

@@

Take a flag variety \((X,G)\) and opposite parabolic subgroups \(P,\op{P}\subseteq G\), so that \(P\) is the stabilizer of \(x_0\in X\).
As above, take their unipotent radicals \(G_+,\op{G_+}\) and the centers \(Z,\op{Z}\) of these.
Let \(\breve{G}\defeq\left<Z,\op{Z}\right>\subseteq G\) be the subgroup generated by \(Z\cup\op{Z}\), \(\breve{P}\defeq\breve{G}\cap P\), \(\breve{X} \defeq \breve{G}/\breve{P}\).

@@

Let \(G'\subseteq G\) be the subgroup leaving \(\breve{X}\subseteq X\) invariant, and \(P':=G'\cap P\).

@@

\breve{P}&

@@

\section{Statement of the main theorem}

@@

With \(\dynkin{A}{o}\) denoting a node which could be either a \(\dynkin{A}{x}\) or \(\dynkin{A}{*}\), the associated cominuscule subvarieties are (allowing some redundacy where it might clarify):

@@

%% A's

@@

\dynkinBrace[p\ge 0]{1}{2}
\dynkinBrace[q\ge 0]{7}{8}

@@

%% B's
\Flag{\dynkin B{x*}}{\dynkin B{x*}}
\Flag{\dynkin B{x**}}{\dynkin B{x**}}

@@

\Flag{\dynkin{B}{ox}}{\dynkin{A}{x}}
\Flag{\dynkin{B}{oxo}}{\dynkin{A}{x}}

@@

\Flag{\dynkin{B}{x*x}}{\dynkin{A}{x*}}
\Flag{\dynkin{B}{x**x}}{\dynkin{A}{x**}}
\begin{dynkinDiagram}{B}{x*.**x}
\dynkinBrace[r-1]{1}{3}
\begin{dynkinDiagram}{B}{x*.**x*}
%\dynkinBrace[r-1]{1}{5}
\dynkinBrace[r-2]{1}{3}

@@

\dynkinBrace[1\le\ell\le r-1]{1}{3}

@@

\Flag{\dynkin{B}{**x}}{\dynkin{A}{**x}}
\Flag{\dynkin{B}{**xo}}{\dynkin{A}{**x}}
\Flag{\dynkin{B}{**xo.oo}}{\dynkin{A}{**x}}
\Flag{\dynkin{B}{***x}}{\dynkin{D}{***x}}
\Flag{\dynkin{B}{***xo}}{\dynkin{D}{***x}}
\LastFlag{12}{B_r}{%

@@

\dynkinBrace[\ell\ge 4]{1}{4}

@@

\Flag{\dynkin C{*x}}{\dynkin C{*x}}
\Flag{\dynkin C{**x}}{\dynkin C{**x}}

@@

\Flag{\dynkin C{xo}}{\dynkin A{x}}
\Flag{\dynkin C{xoo}}{\dynkin A{x}}
\Flag{\dynkin C{xo.ooo}}{\dynkin{A}{x}}%
\Flag{\dynkin C{*xo}}{\dynkin C{*x}}
\Flag{\dynkin C{**xo}}{\dynkin C{**x}}
\Flag{\dynkin C{*.*xo}}{\begin{dynkinDiagram}{C}{**.*x}\dynkinBrace[r-1]{1}{4}\end{dynkinDiagram}}
\LastFlag{7}{C_r}{%

@@

\Flag{\dynkin[parabolic=1]{D}{}}{\dynkin[parabolic=1]{D}{}}
\Flag{\dynkin[parabolic=32]{D}{}}{\dynkin[parabolic=32]{D}{}}
\Flag{\dynkin[parabolic=16]{D}{}}{\dynkin[parabolic=16]{D}{}}
\Flag{\dynkin{D}{oxooooo}}{\dynkin{A}{x}}
\Flag%
\dynkinBrace[3\le\ell\le r-2]{1}{4}

@@

\Flag{\dynkin D{x*.*xx}}{%
\dynkinBrace[r-2]{1}{3}
\Flag%
\begin{dynkinDiagram}{D}{**.*xx}
\begin{dynkinDiagram}{D}{**.**x}
\dynkinLabelRoot{5}{r-1}
\begin{dynkinDiagram}{D}{x*.*x*}
\dynkinBrace[r-1]{1}{4}
\LastFlag{9}{D_r}{%
\begin{dynkinDiagram}{D}{x*.**x}
\dynkinBrace[r-1]{1}{4}
%% E6

@@

%% E7

@@

%% E8

@@

\Flag{\dynkin[ordering=Carter]{E}{*****xxo}}{\dynkin{A}{x*****}}
\Flag{\dynkin[ordering=Carter]{E}{*****x*x}}{\dynkin{A}{x******}}

@@

\Flag{\dynkin[ordering=Carter]{E}{*****x**}}{\dynkin{A}{x*******}}
\LastFlag{10}{E_8}{\dynkin[ordering=Carter]{E}{*******x}}{\dynkin[parabolic=1]{D}{8}}
%% F4

@@

%% G2

@@

For a flag variety \(X=G/P\), the root system is also \emph{\(P\)-graded} by \(\sum n_i\), but summing only over the noncompact (crossed) simple roots \cite{Cap/Slovak:2009} p. 292.
The \(P\)-grade is also called the \emph{\(P\)-height} \cite{Cap/Slovak:2009} p. 292.
A root is \emph{\(P\)-maximal} if it has maximal \(P\)-grade in its irreducible factor.
The \emph{box} is the set of \(P\)-maximal roots, terminology which roughly follows what \cite{Buch.Chaput.Mihalcea.Perrin:2018}, \cite{Lam.Williams:2008} p. 57 might perhaps call the \emph{maximal box}, by analogy with Young tableaux.

@@

\subsection{The box and the Lie algebra}
The unipotent radical \(G_+\subseteq P\) has Lie algebra \(\LieG_+\subseteq\LieP\) the sum of the root vectors of the positively graded roots \cite{Borel:1991} p. 197 Proposition 14.18, \cite{Knapp:2002} p. 482.
The zero graded roots are invariant under reflection in one another.
The sum of the Cartan subalgebra with the sum of the root soaces of the zero graded roots is the semisimple factor of the Levi factor \(\LieGZ\) of \(\LieP\) \cite{Borel:1991} p. 197 Proposition 14.18.
Denote by \(\LieZ\) the Lie algebra of \(Z\).
For any root \(\aa\), denote by \(\gr[\aa]\LieG\) the \(\aa\)-root space of \(\LieG\).
We will prove:
\begin{lemma}\label{lemma:Z.box}
The abelian Lie algebra \(\LieZ\) is the span of the root vectors of the roots of the box:
\LieZ=\bigoplus_{\aa\in\textrm{\textup{box}}}\gr[\aa]\LieG.
Take the box roots (i.e. roots of the box) as vertices of a graph.
If the difference \(\aa-\bb\) of two box roots is a \(P\)-compact simple root,  corresponding to node \(\ell\) of the Dynkin diagram of \((X,G)\), draw an edge from \(\bb\) to \(\aa\) labelled \(\ell\).
Thus the box becomes a graph.
As a graph, the connected components of the box are precisely the boxes of the irreducible factors of the flag variety.
Each component contains the highest root of the factor.
In particular, the box of an irreducible flag variety is connected and contains the highest root.
\subsection{\texorpdfstring{Associated cominuscules in rank \(2\)}{Associated cominuscules in rank 2}}%
\label{subsection:rank.2}

@@

We pick a basis of simple roots in each root system, drawing the simple roots as empty circles:

@@

Each parabolic subgroup has Lie algebra consisting of the sum of the Cartan subalgebra and the root spaces of those roots which lie on or on one side of a hyperplane, so that the compact roots are those on the hyperplane.

@@

We can always pick a basis of simple roots so that every simple root lies on the hyperplane (hence a compact [i.e. uncrossed] simple root) or lies on the chosen side of the hyperplane (a noncompact [i.e. crossed] root):

@@

(We will always pick our hyperplane, for \(G_2\), to allow for the chosen bases of simple roots shown above.)
We can and will pick the hyperplane to be the zero locus of a real linear function taking on value zero on uncrossed simple roots and one on crossed simple roots \cite{Cap/Slovak:2009} p. 239 proposition 3.1.2.
Grade the roots by \(P\)-height, i.e. by sum of coefficients of crossed simple roots.
Our hyperplane is thus always chosen so that we can see the heights, i.e. roots of a given height lie on a parallel hyperplane:

@@

The elements of \(G\) preserving \(\breve{X}\) form a subgroup of \(G\) which maps to \(\breve{G}\), which we call the \emph{automorphism group}.
